---
title: "221108-Part2"
author: "Nichas, Mady"
date: "2022-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA) #keeps out the hashtags in the knits
options(scipen = 999)
```
Murrar and Brauer's (2018) interventions designed to reduce prejudice against Arabs/Muslims. 
Participants (*N* = 193), all who were White, were randomly assigned to one of two conditions where they watched six episodes of the sitcom *Friends* or *Little Mosque on the Prairie*. 
The sitcoms and specific episodes were selected after significant pilot testing.  
The authors felt that both series had characters that were likable and relatable and were engaged in regular activities of daily living. 
The Friends series featured characters who were predominantly White, cis-gendered, and straight. 
The Little Mosque series portrayed the experience of Western Muslims and Arabs as they lived in a small Canadian town. 
This study involved assessment across three waves: baseline (before watching the assigned episodes), post1 (immediately after watching the episodes), and post2 (completed 4-6 weeks after watching the episodes).

The study used feelings and liking thermometers, rating their feelings and liking toward 10 different groups of people on a 0 to 100 sliding scale (with higher scores reflecting greater liking and positive feelings). 
For the purpose of this analysis, the ratings of attitudes toward White people and attitudes toward Arabs/Muslims were used. 
A third metric was introduced by subtracting the attitudes towards Arabs/Muslims from the attitudes toward Whites. 
Higher scores indicated more positive attitudes toward Whites where as low scores indicated no difference in attitudes. 
To recap, there were three potential dependent variables, all continuously scaled:
*AttWhite*: attitudes toward White people; higher scores reflect greater liking
*AttArab*: attitudes toward Arab people; higher scores reflect greater liking
*Diff*: the difference between AttWhite and AttArab; higher scores reflect a greater liking for White people

With random assignment, nearly equal cell sizes, a condition with two levels (Friends, Little Mosque), and three waves (baseline, post1, post2)
Two dependent variables (AttWhite, AttArab)
Wave (baseline, post1, post2) 
COND (condition; Friends, Little_Mosque)
There is also a caseID (repeated three times across the three waves) and rowID (giving each observation within each case an ID)- long-file, where each person has 3 rows of data representing baseline, post1, and post2

#Simulating the data
```{r message=FALSE, library=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
library(tidyverse)
set.seed(210815)
AttWhite <- round(c(rnorm(98, mean = 76.79, sd = 18.55), rnorm(95, mean = 75.37,
    sd = 18.99), rnorm(98, mean = 77.47, sd = 18.95), rnorm(95, mean = 75.81,
    sd = 19.29), rnorm(98, mean = 77.79, sd = 17.25), rnorm(95, mean = 75.89,
    sd = 19.44)), 3)  
#sample size, M and SD for each cell; this will put it in a long file
#set upper bound for variable
AttWhite[AttWhite > 100] <- 100
# set lower bound for variable
AttWhite[AttWhite < 0] <- 0
AttArab <- round(c(rnorm(98, mean = 64.11, sd = 20.97), rnorm(95, mean = 64.37,
    sd = 20.03), rnorm(98, mean = 64.16, sd = 21.64), rnorm(95, mean = 70.52,
    sd = 18.55), rnorm(98, mean = 65.29, sd = 19.76), rnorm(95, mean = 70.3,
    sd = 17.98)), 3)
# set upper bound for variable
AttArab[AttArab > 100] <- 100
# set lower bound for variable
AttArab[AttArab < 0] <- 0
rowID <- factor(seq(1, 579))
caseID <- rep((1:193), 3)
Wave <- c(rep("Baseline", 193), rep("Post1", 193), rep("Post2", 193))
COND <- c(rep("Friends", 98), rep("LittleMosque", 95), rep("Friends", 98),
    rep("LittleMosque", 95), rep("Friends", 98), rep("LittleMosque", 95))
# groups the 3 variables into a single df: ID#, DV, condition
Murrar_df <- data.frame(rowID, caseID, Wave, COND, AttArab, AttWhite)
```

#Time to check the structure
```{r}
str(Murrar_df)
```
#CaseID needs to be a factor
#Wave needs to be an ordered factor
#COND switch to an ordered factor
```{r}
Murrar_df[, "caseID"] <- as.factor(Murrar_df[, "caseID"])

Murrar_df$Wave <- factor(Murrar_df$Wave, levels = c("Baseline", "Post1",
    "Post2"))

Murrar_df$COND <- factor(Murrar_df$COND, levels = c("Friends", "LittleMosque"))
```

#Check structure again
```{r}
str(Murrar_df)
```

#Create new variable for diff between White & Arab scores- preference for white

```{r}
Murrar_df$Diff <- Murrar_df$AttWhite - Murrar_df$AttArab
```

#Getting a look at the overall descriptive stats
```{r}
psych::describe(Murrar_df)
```

#Using the difference score (Diff) as the dependent variable...
```{r}
psych::describeBy(Diff ~ Wave + COND, data = Murrar_df, mat = TRUE)
```
Skew and Kurtosis are below the levels of concern (Kline, 2016)

#Baseline scores for the Friends and Little Mosque conditions are similar
The post1 and post2 difference scores (higher scores indicate more favorable ratings of White individuals) are higher in the Friends condition than in the Little Mosque condition.

#Box plot!
```{r}
CNDwiWV <- ggpubr::ggboxplot(Murrar_df, x = "Wave", y = "Diff", color = "COND",
    palette = "jco", xlab = "Assessment Wave", ylab = "Difference in Attitudes towards Whites and Arabs",
    )
CNDwiWV
```
```{r}
WVwiCND <- ggpubr::ggboxplot(Murrar_df, x = "COND", y = "Diff", color = "Wave",
    palette = "jco", xlab = "Treatment Condition", ylab = "Difference in Attitudes towards Whites and Arabs")
WVwiCND
```
```{r}
Murrar_df %>%
    group_by(Wave, COND) %>%
    rstatix::identify_outliers(Diff)
```

Some outliers but none are extreme
#Shapiro Wilk's- continue to test for assumption of normality
```{r}
Murrar_df %>%
    group_by(Wave, COND) %>%
    rstatix::shapiro_test(Diff)
```
The Shapiro Wilk's test results suggest that the data does not violate the assumption of normality


#QQ plots.


```{r}
ggpubr::ggqqplot(Murrar_df, "Diff", ggtheme = theme_bw()) + facet_grid(Wave ~
    COND)
```


Levene's test to check for the assumption of homogeneity

```{r}
Murrar_df %>%
    group_by(Wave) %>%
    rstatix::levene_test(Diff ~ COND)
```
Levene's test has indicated no violation of the homogeneity of variance
Friends and Little Mosque conditions at 
Baseline $(*F* [1, 191] = 0.050, *p* = .834)$
Post1 $(*F* [1, 191] = 0.439, *p* = .508)$
Post2 $(*F* [1, 191] = 1.533, *p* = .217)$


Box's M test evaluates if two or more covariance matrices are homogeneous. 
(low power in small sample sizes and is overly sensitive in large sample sizes)

```{r}
rstatix::box_m(Murrar_df[, "Diff", drop = FALSE], Murrar_df$COND)
```
Box's M indicated no violation of the homogeneity of covariance matrices assumption $(*M* = .111, *p* = .73)$

```{r}
Diff_2way <- rstatix::anova_test(
  data = Murrar_df, dv = Diff, wid = caseID, 
  between = COND, within = Wave 
  )
Diff_2way
```
Mauchly's test indicated no violation of the sphericity assumption for the main effect $(*W* = 0.989, *p* = .346)$ and interaction effect $(*W* = 0.989, *p* = .346)$. 

Results of the omnibus ANOVA indicated nonsignificant effects for all conditions 
Main effect for condition $(*F*[1, 191] = 3.749, *p* = .054, \eta^{2} = 0.006)$
Main effect for wave $(*F*[2, 382] = 0.939, *p* = .392, \eta^{2} = 0.003)$
Interaction effect $(*F*[2, 382] = 1.133, *p* = 0.323, \eta^{2} = 0.004)$

We're here to practice...
```{r}
# Three one-way ANOVAs for each of the waves, separately the between = Cond means that each level of cond will be compared - 'bonferroni' gets us both the standard and adjusted p- values
SimpleWave <- Murrar_df %>%
    group_by(Wave) %>%
    rstatix::anova_test(dv = Diff, wid = caseID, between = COND) %>%
    rstatix::get_anova_table() %>%
    rstatix::adjust_pvalue(method = "bonferroni")
SimpleWave
```

Pre: $*F* (1, 191) = 0.000, *p* = .989, \eta^{2} = 0.000$ 
Post1: $*F* (1, 191) = 1.364, *p* = .244, \eta^{2} = 0.007$ 
Post2: $*F* (1, 191) = 4.606, *p* = .033*, \eta^{2} = 0.024$ (a small effect size)
(Interpretation for the eta-squared are .01 ~ small, .06 ~ medium, and >.14 ~ large)

Because there are only two levels (Friends, Little Mosque) within each wave (baseline, post1, post2), this simple effects analysis is complete with the three pairwise comparisons. 

Type I error- traditional Bonferroni, we would use $\alpha$ = .017 (.05/3).

```{r}
.05/3
```

```{r}
SimpleCond <- Murrar_df %>%
    group_by(COND) %>%
    rstatix::anova_test(dv = Diff, wid = caseID, within = Wave) %>%
    rstatix::get_anova_table() %>%
    rstatix::adjust_pvalue(method = "bonferroni")
SimpleCond
```

Friends: $*F*(2, 194) = 1.107, *p* = 0.333, \eta^{2} = 0.007$ 
Little Mosque: $*F* (2, 188) = 0.973, *p* = 0.380, \eta^{2} = 0.008$ 

#Pairwise comparisons

```{r}
pwcWVwiGP <- Murrar_df %>%
    group_by(COND) %>%
    rstatix::pairwise_t_test(Diff ~ Wave, paired = TRUE, detailed = TRUE,
        p.adjust.method = "bonferroni")  #%>%
# select(-df, -statistic, -p) # Remove details
pwcWVwiGP
```

Three more pairwise comparisons in the Little Mosque condition
Divide .05/3 and interpret these pairwise comparisons with an alpha of .017

We find a non-significant differences between all
We followed the nonsignificant interaction effect with an evaluation of simple main effects of wave within condition. There were non-significant difference within the the Friends condition $*F*(2, 194) = 1.107, *p* = 0.333, \eta^{2} = 0.007$ 
There were nonsignificant differences in the Little Mosque condition $*F* (2, 188) = 0.973, *p* = 0.380, \eta^{2} = 0.008$ 
We would follow up the significant simple main effect for the Little Mosque condition with pairwiwse comparisons... we're here to practice
At this level we controlled for Type I error by dividing alpha (.05) by the number of paired comparisons (3). We did not find a significant difference between baseline, post1, or post2 

```{r}
Murrar_df %>%
    rstatix::pairwise_t_test(Diff ~ Wave, paired = TRUE, p.adjust.method = "bonferroni")
```

```{r}
pwcWVwiGP <- pwcWVwiGP %>%
    rstatix::add_xy_position(x = "Wave")
CNDwiWV + ggpubr::stat_pvalue_manual(pwcWVwiGP, tip.length = 0, hide.ns = TRUE) +
    labs(subtitle = rstatix::get_test_label(Diff_2way, detailed = TRUE),
        caption = rstatix::get_pwc_label(pwcWVwiGP))
```

```{r}
# pwcWVwiGP were my pairwise comparisons for the simple effect
# UE_2way was my omnibus ANOVA object WVwiCND was the boxplot before
# I did the ANOVA
pwcWVwiGP <- pwcWVwiGP %>%
    rstatix::add_xy_position(x = "Wave")
WVwiCND + ggpubr::stat_pvalue_manual(pwcWVwiGP, tip.length = 0, hide.ns = TRUE) +
    labs(subtitle = rstatix::get_test_label(Diff_2way, detailed = TRUE),
        caption = rstatix::get_pwc_label(pwcWVwiGP))
```

Effect size
```{r}
effectsize::eta2_to_f(0.017) 
```
```{r}
WebPower::wp.rmanova(n=193, ng=2, nm=3, f = .1315, nscor = .99, alpha = .05, power = NULL, type = 2)
```
35% power

```{r}
WebPower::wp.rmanova(n = NULL, ng = 2, nm = 3, f = 0.1315, nscor = 0.99,
    alpha = 0.05, power = 0.8, type = 2)
```

Need n of 563