---
title: "Problem3-MiDataEsSuData"
author: "Nichas, Mady"
date: "2022-09-21"
output: html_document
---
```{r}
library(tidyverse)
set.seed(123)
Blue_mu <- c(1.52, 1.72, 2.69, 1.71, 2.14, 2.35, 2.42)
Blue_stddev <- c(2.52, 2.04, 0.47, 0.7, 0.8, 2.41, 3.36)
Blue_corMat <- matrix(c(1, 0.69, 0.19, 0.28, 0.32, 0.08, 0.23, 0.69, 1,
    0.2, 0.29, 0.33, 0.13, 0.25, 0.19, 0.2, 1, 0.5, 0.5, -0.04, 0.09, 0.28,
    0.29, 0.5, 1, 0.76, 0.04, 0.18, 0.32, 0.33, 0.5, 0.76, 1, 0.1, 0.21,
    0.08, 0.13, -0.04, 0.04, 0.1, 1, 0.42, 0.23, 0.25, 0.09, 0.18, 0.21,
    0.42, 1), ncol = 7)
```


```{r}
Blue_covMat <- Blue_stddev %*% t(Blue_stddev) * Blue_corMat

Blue_dat <- MASS::mvrnorm(n = 389, mu = Blue_mu, Sigma = Blue_covMat,empirical = TRUE)
```
```{r}
 Blue_df <- as.data.frame(Blue_dat)
```
```{r}
library(tidyverse)
Blue_df <- rename(Blue_df, Sun = V1, Mon = V2, Tues = V3, Wed = V4, Thurs = V5, Fri = V6, Sat = V7)
```
```{r}
Blue_df$Sun[Blue_df$Sun > 24] <- 24
Blue_df$Sun[Blue_df$Sun < 0] <- 0

Blue_df$Mon[Blue_df$Mon > 24] <- 24
Blue_df$Mon[Blue_df$Mon < 0] <- 0

Blue_df$Tues[Blue_df$Tues > 24] <- 24
Blue_df$Tues[Blue_df$Tues < 0] <- 0

Blue_df$Wed[Blue_df$Wed > 24] <- 24
Blue_df$Wed[Blue_df$Wed < 0] <- 0

Blue_df$Thurs[Blue_df$Thurs > 24] <- 24
Blue_df$Thurs[Blue_df$Thurs < 0] <- 0

Blue_df$Fri[Blue_df$Fri > 24] <- 24
Blue_df$Fri[Blue_df$Fri < 0] <- 

Blue_df$Sat[Blue_df$Sat > 24] <- 24
Blue_df$Sat[Blue_df$Sat < 0] <- 0

Blue_df$BandanaColor <- "Blue"
```


```{r}
Pink_mu <- c(4.45, 3.84, 2.6, 1.84, 2.1, 2.81, 2.14)
Pink_stddev <- c(4.22, 3.08, 0.89, 0.8, 0.81, 2.49, 3.24)
Pink_corMat <- matrix(c(1, 0.81, 0.17, 0.15, 0.09, 0.05, -0.16, 0.81,
    1, 0.17, 0.21, 0.11, 0.09, -0.01, 0.17, 0.17, 1, 0.59, 0.54, 0.05,
    0.24, 0.15, 0.21, 0.59, 1, 0.72, 0.12, 0.22, 0.09, 0.11, 0.54, 0.72,
    1, 0.21, 0.4, 0.05, 0.09, 0.05, 0.12, 0.21, 1, 0.65, -0.16, -0.01,
    0.24, 0.22, 0.4, 0.65, 1), ncol = 7)
Pink_covMat <- Pink_stddev %*% t(Pink_stddev) * Pink_corMat
Pink_dat <- MASS::mvrnorm(n = 133, mu = Pink_mu, Sigma = Pink_covMat,
    empirical = TRUE)
Pink_df <- as.data.frame(Pink_dat)
Pink_df <- rename(Pink_df, Sun = V1, Mon = V2, Tues = V3, Wed = V4,
    Thurs = V5, Fri = V6, Sat = V7)
```


```{r}

# set upper and lower bound for each variable
Pink_df$Sun[Pink_df$Sun > 24] <- 24
Pink_df$Sun[Pink_df$Sun < 0] <- 0

Pink_df$Mon[Pink_df$Mon > 24] <- 24
Pink_df$Mon[Pink_df$Mon < 0] <- 0

Pink_df$Tues[Pink_df$Tues > 24] <- 24
Pink_df$Tues[Pink_df$Tues < 0] <- 0

Pink_df$Wed[Pink_df$Wed > 24] <- 24
Pink_df$Wed[Pink_df$Wed < 0] <- 0

Pink_df$Thurs[Pink_df$Thurs > 24] <- 24
Pink_df$Thurs[Pink_df$Thurs < 0] <- 0

Pink_df$Fri[Pink_df$Fri > 24] <- 24
Pink_df$Fri[Pink_df$Fri < 0] <- 0

Pink_df$Sat[Pink_df$Sat > 24] <- 
Pink_df$Sat[Pink_df$Sat < 0] <- 0

Pink_df$BandanaColor <- "Pink"

```
```{r}
Purple_mu <- c(1.56, 2.34, 2.69, 1.81, 2.17, 3.47, 2.69)
Purple_stddev <- c(2.46, 2.49, 0.86, 0.71, 0.78, 2.59, 3.76)
Purple_corMat <- matrix(c(1, 0.78, 0.27, 0.36, 0.42, -0.06, 0.08, 0.78,
    1, 0.33, 0.26, 0.35, -0.11, -0.02, 0.27, 0.33, 1, 0.62, 0.64, -0.04,
    0.15, 0.36, 0.26, 0.62, 1, 0.81, -0.08, 0.17, 0.42, 0.35, 0.64, 0.81,
    1, -0.06, 0.15, -0.06, -0.11, -0.04, -0.08, -0.06, 1, 0.6, 0.08, -0.02,
    0.15, 0.17, 0.15, 0.6, 1), ncol = 7)
Purple_covMat <- Purple_stddev %*% t(Purple_stddev) * Purple_corMat
Purple_dat <- MASS::mvrnorm(n = 182, mu = Purple_mu, Sigma = Purple_covMat,
    empirical = TRUE)
Purple_df <- as.data.frame(Purple_dat)
Purple_df <- rename(Purple_df, Sun = V1, Mon = V2, Tues = V3, Wed = V4,
    Thurs = V5, Fri = V6, Sat = V7)
Purple_df$Sun[Purple_df$Sun > 24] <- 24
Purple_df$Sun[Purple_df$Sun < 0] <- 0

Purple_df$Mon[Purple_df$Mon > 24] <- 24
Purple_df$Mon[Purple_df$Mon < 0] <- 0

Purple_df$Tues[Purple_df$Tues > 24] <- 24
Purple_df$Tues[Purple_df$Tues < 0] <- 0

Purple_df$Wed[Purple_df$Wed > 24] <- 24
Purple_df$Wed[Purple_df$Wed < 0] <- 0

Purple_df$Thurs[Purple_df$Thurs > 24] <- 24
Purple_df$Thurs[Purple_df$Thurs < 0] <- 0

Purple_df$Fri[Purple_df$Fri > 24] <- 24
Purple_df$Fri[Purple_df$Fri < 0] <- 0

Purple_df$Sat[Purple_df$Sat > 24] <- 24
Purple_df$Sat[Purple_df$Sat < 0] <- 0

Purple_df$BandanaColor <- "Purple"

```
```{r}
Bandana_sim_df <- bind_rows(Blue_df, Pink_df, Purple_df)
```
```{r}
saveRDS(Bandana_sim_df, 'Bandana_RDS.rds')
```
```{r}
df <- readRDS('Bandana_RDS.rds')
```
```{r}
str(df)
```
```{r}
library(tidyverse)
df <- df %>%
    dplyr::mutate(BandanaColor = as.factor(BandanaColor))
```
```{r}
str(df)
```

```{r}
ggpubr::gghistogram(df$Sun, xlab="Negative Affect", ylab = "Frequency", add = "mean", rug=TRUE, color = "#00BFBB")

```
```{r}
Blue <- c(1.52, 1.52, 2.69, 1.71, 2.14, 2.35, 2.42)
Blue <- as.data.frame(Blue)
mean(Blue$Blue, na.rm = TRUE)
```
```{r}
median(Blue$Blue, na.rm = TRUE)

```

```{r}
lsr::modeOf(Blue$Blue)
```
```{r}
ggpubr::gghistogram(df$Sat, xlab="Negative Affect", ylab = "Frequency", add = "mean", rug=TRUE, color = "#00AFBB")
```
```{r}
min(Blue$Blue, na.rm = TRUE)
max(Blue$Blue, na.rm = TRUE)
range(Blue$Blue, na.rm = TRUE)
```

```{r}
min(df$Mon, na.rm = TRUE)
max(df$Mon, na.rm = TRUE)
range(df$Mon, na.rm = TRUE)
```
```{r}
quantile(df$Mon, probs = c(0.25, 0.75))
```

```{r}
3.73-.53
```

```{r}
IQR(df$Mon, na.rm = TRUE)
```

```{r}
Blue$mean <- mean(Blue$Blue, na.rm = TRUE)
head(Blue)
```

```{r}
Blue$mdev <- Blue$Blue - Blue$mean
head(Blue)
```
```{r}
round(sum(Blue$mdev, na.rm = TRUE), 3)
round(mean(Blue$mdev, na.rm = TRUE), 3)
```

```{r}
Blue$abslt_m <- abs(Blue$mdev)
head(Blue)
```
```{r}
round(mean(Blue$abslt_m, na.rm = TRUE), 3)
```
```{r}
library(tidyverse)
df_3vars <- df %>%
    dplyr::select(Tues, Wed, Thurs)
```
```{r}
psych::describe(df_3vars)
```
```{r}
library(tidyverse)

df4corr <- df %>%
    dplyr::select(Tues, Wed)

head(df4corr)
```
```{r}
df4corr$MDTues <- df4corr$Tues - mean(df4corr$Tues)

df4corr$MDWed <- df4corr$Wed - mean(df4corr$Wed)

head(df4corr)
```
```{r}
df4corr$crossproductXY <- df4corr$MDTues * df4corr$MDWed
head(df4corr)
sum(df4corr$crossproductXY)
```

```{r}
cov <- 1/(nrow(df4corr) - 1) * sum(df4corr$crossproductXY)
cov
```
```{r}
cov/(sd(df4corr$Tues) * sd(df4corr$Wed))
```

```{r}
cor.test(df4corr$Tues, df4corr$Wed)
```

```{r}
psych::pairs.panels(df_3vars, pch = ".")
```
```{r}
Table1_Cor <- apaTables::apa.cor.table(df_3vars, filename = "Table1_Cor.doc",
    table.number = 1, show.conf.interval = FALSE, landscape = TRUE)
```

```{r}
print(Table1_Cor)

```
Preliminary Results
Our sample included 704 observations identified by the day of the week (Sun, Mon, Tues, Wed, Thurs, Fri, Sat) and amount of time each color bandana was worn. Visual inspection of the three variables of interest (Tues, Wed, Thurs) combined with formal evaluation of skewness and kurtosis suggested that their distributions did not violate the assumption of univariate normality. Means, standard deviations, and a correlation matrix are presented in Table 1. We noted that the correlation between Tues and Thurs was strong (r = 0.76); correlations between remaining variables were smaller but still moderate.
